Using vector space models based on LDA and LSA for automatically grading exam questions

### 2018-06-15 tex/article.tex
Minor changes to the discussion.

%\subsection{Best-performing individual models}
%As explained in Section \ref{sec:averageoutcomes}, all cells in Table \ref{table1} represent average outcomes.

@@

Looking back on the literature, \citeasnoun{kakkonen2008} asked whether their LSA models perhaps outperformed their LDA models because they were trained on a small set of student answers only. My results support this explanation: when training on the student answers LSA outperformed LDA, but it was the other way around when using a more diverse document collection (i.e., sentences in a psychology textbook chapter). This raises the question as to why the two types of training data affected the topic models differently (the size of the training sets was similar).

One explanation can be sought in the optimal number of topics for the LDA models that resulted from a grid search (see 2.3.3). This number was different for the two types of training data. In the student answers, the grid search had shown that the optimal number of topics was 2. This is the lowest possible number of topics that an LDA model can have. This outcome makes sense because in essence the student answers all revolved around the same topic (i.e., Whorf's language theory). As a result, however, the resulting vector space models only had two dimensions, which does not allow very fine distinctions between answers of different qualities. When visually inspecting the scatter plots of the outcomes of these models (not printed here), it can be seen that these models predict a 10 for almost all students, resulting in very low correlations. Thus, training on a more diverse corpus such as a textbook seems more suitable when working with LDA models. Indeed, Table 1 shows better results for LDA models that were trained on the textbook and used 20 topics, as compared to LDA models that were trained on student answers.

>>>>

@@

Looking back on the literature, \citeasnoun{kakkonen2008} asked whether their LSA models perhaps outperformed their LDA models because they were trained on a small set of student answers only. My results support this explanation: when training on the student answers LSA outperformed LDA, but it was the other way around when using a more diverse document collection (i.e., sentences in a psychology textbook chapter). This raises the question as to why the two types of training data affected the topic models differently (while the size of the training sets was similar).

One explanation can be sought in the optimal number of topics for the LDA models that resulted from a grid search (see Section \ref{sec:training}). This number was different for the two types of training data. In the student answers, the grid search had shown that the optimal number of topics was 2. This is the lowest possible number of topics that an LDA model can have. This outcome makes sense because in essence the student answers all revolved around the same topic (i.e., Whorf's language theory). As a result, however, the resulting vector space models only had two dimensions, which does not allow very fine distinctions between answers of different qualities. When visually inspecting the scatter plots of the outcomes of these models (not printed here), it can be seen that these models predict a 10 for almost all students, resulting in very low correlations. Thus, training on a more diverse corpus such as a textbook seems more suitable when working with LDA models. Indeed, Table 1 shows better results for LDA models that were trained on the textbook and used 20 topics, as compared to LDA models that were trained on student answers.
